Security Properties for Stack Safety

### 2021-05-14 short/main.tex
Polishing on the first part

The call stack is a perennial target for low-level attacks, leading to
consequences ranging from leakage or corruption of private stack data to
control-flow hijacking. A profusion of
software and hardware protections have been proposed for detecting or
preventing such attacks,
%
including stack canaries~\citep{Cowan+98},
bounds checking~\citep{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
split stacks~\citep{Kuznetsov+14},
shadow stacks~\citep{Dang+15,Shanbhogue+19},
capabilities~\citep{Woodruff+14,Chisnall+15,SkorstengaardLocal,SkorstengaardSTK,Georges+21},
and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}.
%
The protections offered by such mechanisms are commonly described in terms
of concrete examples of attacks that they can prevent.
At best, they define stack safety by reference to an idealized machine that
is arguably stack safe by construction \citep{SkorstengaardSTK}.
But these mechanisms can be intricate, and it would be useful to have a precise, generic, and formal
specification for stack safety, both to compare the security claims of different
enforcement techniques and to rigorously validate such claims.

We propose such a characterization,
using the technical framework of language-based security.
Stack safety protects a caller from its callee, guaranteeing
the {\em integrity} and {\em confidentiality} of the caller's local state
until it regains control.

We formalize integrity and confidentiality as trace (hyper-)properties, in two different
variants: {\em stepwise} variants, in which a caller's data must {\em never} be
read or modified during a call, and {\em observational} ones, in which
callees may read from and write to their caller's stack frame, as
long as these ``risky'' behaviors do not affect the system's observable
behavior. The observational properties are more extensional, and
any reasonable protection mechanism ought to enforce them,
even if it does not prevent every single dangerous read or write.

Confidentiality is especially interesting, as it is based on the traditional
notion of noninterference. But where noninterference is normally presented as an
end-to-end hyper-property, stack confidentiality is noninterference applied over
multiple nested subtraces -- each individual call's behavior must be invariant
regardless of the state of the stack at its entry.

>>>>

The call stack is a perennial target for low-level attacks, leading to
consequences  from leakage or corruption of private stack data to
control-flow hijacking. A profusion of
software and hardware protections have been proposed for detecting or
preventing such attacks,
%
including stack canaries~\citep{Cowan+98},
bounds checking~\citep{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
split stacks~\citep{Kuznetsov+14},
shadow stacks~\citep{Dang+15,Shanbhogue+19},
capabilities~\citep{Woodruff+14,Chisnall+15,SkorstengaardLocal,SkorstengaardSTK,Georges+21},
and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}.
%
The protections offered by such mechanisms are commonly described in terms
of concrete examples of attacks that they can prevent.
At best, they define stack safety by reference to an idealized machine that
is arguably stack safe by construction \citep{SkorstengaardSTK}.
But these mechanisms can be intricate, and it would be useful to have a precise, generic, and formal
specification for stack safety, both to compare the security claims of different
enforcement techniques and to rigorously validate such claims.

We propose such a characterization
using the tools of language-based security: The informal claim that
``stack safety protects a caller from its callee'' amounts to saying
that it guarantees
the {\em integrity} and {\em confidentiality} of the caller's local state
until it regains control.

We formalize these integrity and confidentiality requirements as trace
(hyper-)properties, with two different
variants: {\em stepwise} variants, in which a caller's data must {\em never} be
read or modified during a call, and {\em observational} ones, in which
callees may read from and write to their caller's stack frame, as
long as these ``risky'' behaviors do not affect the system's observable
behavior. The observational properties are more extensional and thus
allow ``lazier'' protection mechanism that permit ``dangerous'' reads or
writes as long as they do not affect the program's visible outcomes.

Confidentiality is especially interesting.  It is based on a traditional
notion of noninterference, but where ordinary noninterference is 
an end-to-end hyper-property on whole program runs, stack confidentiality is
a ``nested'' form of 
noninterference applied to
subtraces of the whole program trace, requiring that, between each call
instruction and the corresponding return, the
caller's stack frame is protected from the callee.
